\begin{abstract}
Federated IoT anomaly detectors often train one shared autoencoder and then apply one shared reconstruction-error threshold. When physical devices produce different benign traffic patterns, one threshold can concentrate false alarms on a subset of clients. This paper presents a controlled threshold-calibration study---not a new detection algorithm---that isolates threshold scope as the only experimental variable after federated training under a frequently synchronized FedAvg encoder ($E{=}1$, full client participation): the same autoencoder, seeds, and score artifacts are reused while shared, per-client, family-based, and cluster-based thresholds are compared.

On N-BaIoT, where each client is a physical IoT device, per-client calibration (B2) reduces the coefficient of variation of false-positive rate from 1.017 to 0.299 ($\Delta = 0.718$, all five seed-level deltas positive). Cluster-based calibration (B4) recovers approximately 52\% of this improvement without requiring a device taxonomy. The gain is not a uniform detection-quality improvement: CV(TPR) increases slightly and the lower-tail Macro-F1 degradation is concentrated in a single low-separability client (Ennio Doorbell), representing a client-specific failure mode rather than a universal tradeoff. On CICIoT2023, where randomized file-level pseudo-clients (not physical IoT devices) are near-homogeneous, per-client calibration offers no improvement under this randomized partition; chronological or physical-device CICIoT2023 partitions remain untested. A Dirichlet severity sweep supports the interpretation that threshold personalization is useful mainly under stronger client heterogeneity.
\end{abstract}

\begin{IEEEkeywords}
federated learning, IoT anomaly detection, threshold calibration, false-positive-rate dispersion, autoencoder, non-IID data, personalization.
\end{IEEEkeywords}
