\section{Conclusion}
\label{sec:conclusion}

This study holds the FedAvg AE ($E{=}1$, full participation), training protocol, and seeds fixed across B1--B4;
only the post-training threshold calibration scope varies.
On heterogeneous N-BaIoT devices, per-client calibration (B2) provides the strongest
FPR-dispersion reduction ($\Delta = 0.718$, all five seed deltas positive, range $[0.588,\,0.772]$);
cluster-mean calibration (B4) recovers a meaningful but partial share
($\approx$52\% of B2's improvement) without requiring a device taxonomy, at the cost
of transmitting distributional metadata that is not private.
Family-mean calibration (B3) performs poorly: coarse device-type labels do not align
with reconstruction-error calibration structure.

When benign error distributions are near-homogeneous (CICIoT2023 under our file-level
pseudo-client partition, Dirichlet near-IID), no dispersion reduction is observed---this is
an applicability boundary rather than evidence that personalization is unnecessary for these datasets generally.
The main tradeoff under heterogeneous conditions is that B2 reduces FPR dispersion
and improves mean Macro-F1, but the lower-tail P10 Macro-F1 degradation is concentrated
in the low-separability Ennio Doorbell client---a client-specific failure of percentile
calibration rather than evidence of a universal FPR--TPR tradeoff.
Per-client calibration is an FPR-equity intervention, not a uniform detection-quality improvement.

\textbf{Practitioner guidance.}
First, measure benign calibration-error heterogeneity across clients, for example
using pairwise distribution divergence or dispersion of local calibration thresholds.
Second, choose the threshold policy according to that heterogeneity and deployment
constraints: use B1 when clients are near-homogeneous or per-client state is
undesirable; use B2 when local calibration is available and false-alarm equity is the
priority; use B4 when taxonomy-free grouping is acceptable, while accounting for the
metadata leakage introduced by shared distributional fingerprints.

\textbf{Limitations.}
Regime~A uses only 9 physical devices (B4 is therefore exploratory at this scale);
Regime~B uses randomized file-level pseudo-clients without temporal ordering, not
physical IoT devices; B4 fingerprints leak distributional metadata;
no adversarial robustness, formal privacy, or real hardware evaluation is provided;
five seeds limit statistical power (seed-level descriptive summaries are reported rather than confidence intervals).
Because $E{=}1$ and all clients participate in every round, local weight drift is
intentionally limited; evaluating whether the same threshold-personalization gains persist
under larger local-epoch settings (which increase client drift and alter reconstruction-error
geometry) is left as a sensitivity direction.
A physical-device or attack-source CICIoT2023 partition would be required to test DATP under
stronger CICIoT2023 heterogeneity; such a partition is outside the present controlled study.
Threshold personalization is complementary to---not a replacement for---model-level personalization.

\textbf{Future work:} larger multi-vendor physical fleets,
alternative CICIoT2023 partitions (physical-device identity, attack-source, attack-type),
sensitivity to larger local-epoch settings ($E > 1$),
privacy-preserving grouping
(secure aggregation or differential privacy for fingerprint transmission),
model-level personalization comparison, adversarial robustness under poisoning
and evasion, and deployment validation under concept drift.

\noindent\textbf{Availability.}
A replication package containing source code, configuration files, per-seed results,
and table/figure generation scripts is archived on Zenodo (DOI: 10.5281/zenodo.20315654).
Detailed reproduction commands are provided in the repository documentation.
