\section{Experimental Design}
\label{sec:experiments}

\subsection{Datasets and Regimes}

\textbf{Regime A — N-BaIoT (confirmatory).}
The N-BaIoT dataset~\cite{meidan2018n} contains traffic from nine physical IoT devices
grouped by the configured device-family taxonomy, infected with Mirai and BASHLITE botnets.
Each physical device is one FL client ($K=9$).
The N-BaIoT natural partition mixes device-specific benign feature heterogeneity with some
attack-variant skew, and is therefore treated as a realistic natural partition rather than a
pure feature-skew isolation regime.
Regime A is the primary planned condition for RQ2.

\textbf{Regime B --- CICIoT2023 (file-level applicability check).}
CICIoT2023~\cite{neto2023ciciot2023} provides 63 file-defined pseudo-clients under the partition used here.
Regime~B uses randomized file-level pseudo-clients, not physical IoT devices.
Each of the 63 file-defined clients corresponds to one merged flow-level traffic file from the dataset's
\texttt{MERGED\_CSV} directory; client boundaries are defined by the dataset's own
file-level granularity, not by physical device identity.
Within each client, benign records are randomly shuffled (no temporal ordering is preserved)
and split 70/15/15 into train, calibration, and test-benign partitions; attack records
are reserved exclusively for test-attack evaluation.
A per-client cap of 50{,}000 total records is applied, reserving up to 10{,}000 attack records;
this bounds computation and balances benign/attack evaluation volume while preserving
the fixed controlled comparison (after capping, the effective per-client evaluation mix is
approximately 20\% benign and 80\% attack, with median benign count ${\approx}2{,}500$
per client; exact counts are recorded in the replication artifact).
Regime~B tests whether the Regime~A personalization benefit holds when benign
reconstruction-error distributions are near-homogeneous under this partition;
it is not part of the primary B1--B4 comparison.

\textbf{Regime C — Dirichlet non-IID severity (controlled sweep).}
N-BaIoT traffic is repartitioned using a Dirichlet distribution with
$\alpha \in \{0.1,\, 0.3,\, 0.5,\, 1.0,\, 10.0,\, \mathrm{IID}\}$ across $K=20$ synthetic clients.
Lower $\alpha$ means higher heterogeneity.
Regime C provides secondary controlled non-IID severity evidence for the conditions under which threshold personalization is beneficial.

\subsection{Architecture and Training}
The AE encoder has dimensions $[d, 80, 40, 20]$ (ReLU activations, no batch normalization),
where $d=115$ for N-BaIoT and $d=39$ for CICIoT2023.
FL simulation uses the Flower framework~\cite{beutel2020flower} with FedAvg
for up to 150 rounds and a relative-change convergence criterion
(window~$= 10$ rounds, tolerance~$= 0.005$); clients perform $E{=}1$ local epoch
per round with Adam ($\text{lr}=10^{-3}$, batch size 256).
The choice $E{=}1$ keeps training dynamics conservative and fixed while isolating
post-training threshold scope; evaluating larger $E$ is future work because it changes
client drift and reconstruction-error geometry.
With $E{=}1$ and full client participation, FedAvg approaches synchronous distributed
training with frequent aggregation; the FedAvg framing is retained for consistency with
FL-AE literature, and the post-training threshold comparison is invariant to this distinction.
B0 trains a centralized AE on pooled benign data, applying the pooled 95th-percentile
threshold; it is excluded from the B1--B4 controlled ladder.

\subsection{Evaluation Protocol}
Five seeds (0--4) are used. The AE is trained once per seed; B1--B4 thresholds
are derived from shared per-client score artifacts.
The calibration quantile is $q=0.95$;
clients with fewer than $n_{\min}=100$ benign calibration samples are calibration-pending,
receive $\tau_{\text{global}}$ as fallback, and are excluded from CV(FPR).
Budget-capped runs (150 rounds without satisfying the convergence criterion) reuse
the final AE across all B1--B4 derivations, preserving the threshold-only
comparison invariant.
In Regime~A all 5 seeds converged (rounds 51--93);
in Regime~B all 5 were budget-capped;
in Regime~C, 21 of 30 seeds converged (9 budget-capped, concentrated at $\alpha\in\{10.0,\mathrm{IID}\}$).

\subsection{Statistical Analysis}
The primary statistic is the per-seed delta $\Delta_s = \text{CV(FPR)}_{\text{B1},s} - \text{CV(FPR)}_{\text{B2},s}$.
With only five seeds, seed-level descriptive summaries---mean, standard deviation,
range, and sign consistency---are the primary reporting mechanism.
A bootstrap CI (10{,}000 resamples of the five seed-level deltas) is reported as a
supplementary interval estimate; it excludes zero but should be interpreted cautiously
given the small seed count.
Coverage ratio $\rho_{\text{cov}} = |\mathcal{K}_{\mathrm{elig}}| / |\mathcal{K}|$ is reported for all conditions.

Table~\ref{tab:regimes} summarizes the three regimes.
Table~\ref{tab:repro} provides the full reproducibility specification.

\textbf{Environment.}
All experiments ran on a single workstation: AMD Ryzen~5 7600 (6-core), 12\,GB RAM,
NVIDIA GeForce RTX~5060~Ti (16\,GB), Ubuntu~24.04 LTS.
Software: Python~3.12.3, PyTorch~2.11.0 (CUDA~13.0), Flower~1.29.0,
Ray~2.55.0, NumPy~2.4.4, pandas~2.3.3, scikit-learn~1.8.0.
Runtime is not used as an evaluation endpoint in this threshold-scope study.

\begin{table}[t]
  \caption{Experimental Regime Summary}
  \label{tab:regimes}
  \centering
  \footnotesize
  \renewcommand{\arraystretch}{1.2}%
  \begin{tabular}{@{}llcl@{}}
    \toprule
    Regime & Dataset & $K$ & Purpose \\
    \midrule
    A & N-BaIoT & 9 & Confirmatory (B1 vs B2) \\
    B & CICIoT2023 & 63 & Condition-of-applicability check \\
    C & N-BaIoT (Dirichlet) & 20 & Non-IID severity sweep (secondary) \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[t]
  \caption{Reproducibility specification.}
  \label{tab:repro}
  \centering
  \footnotesize
  \renewcommand{\arraystretch}{1.15}%
  \setlength{\tabcolsep}{4pt}%
  \begin{tabular}{@{}lp{0.55\columnwidth}@{}}
    \toprule
    Parameter & Value \\
    \midrule
    AE encoder dims & $[d, 80, 40, 20]$ \\
    Loss / Optimizer & MSE / Adam (lr $10^{-3}$) \\
    Batch size & 256 \\
    Local epochs per round & 1 \\
    Rounds budget & 150 (convergence early-stop) \\
    Convergence criterion & rel.\ change $< 0.005$, window 10 \\
    Client participation & Full (all clients each round) \\
    Normalization scope & Per-client (StandardScaler) \\
    Seeds & $\{0, 1, 2, 3, 4\}$ \\
    Threshold quantile $q$ & 0.95 \\
    Cal.\ eligibility $n_{\min}$ & 100 benign cal.\ samples \\
    N-BaIoT split ratios & 60/20/18 (train/cal/test, gapped)$^{a}$ \\
    CICIoT2023 split & 70/15/15 benign; attack test-only$^{b}$ \\
    CICIoT2023 cap & 50{,}000 total per client (10{,}000 attack) \\
    Score artifact reuse & Train once; B1--B4 from shared scores \\
    \bottomrule
  \end{tabular}
  \par\vspace{2pt}
  \parbox{\columnwidth}{\footnotesize
    $^{a}$ Chronological split with two 1\%-wide temporal gaps to prevent leakage.
    Nominal test fraction~$= 0.18$; gap buffers excluded from all partitions.\\[1pt]
    $^{b}$ CICIoT2023: benign split 70/15/15; attack reserved for test only
    (\texttt{calibration\_benign\_only=True}).
  }
  \vspace{-2pt}
\end{table}
